---
title: "Importance Values and Diversity Analysis"
output: html_notebook
---

Going over Leslie's data to see if I get the same results she does.

Prep the workspace
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
```

Start with one file. It has Leslie's data plus the data from the relevant FE forests.

```{r}
DF<-read.csv("Data/forests.csv")
```

Plot some data and look for issues
```{r}
ggplot(DF, aes(DBH))+
   geom_histogram()
```
Looks like there might be a really big tree or two.  Check them out. 

```{r}
big<-filter(DF, DBH >= 60)
```
Two big sugar maples at donnerville.  Reasonable.


Now look at species to see if they are all named correctly. Start by making some factor variables
```{r}
species<-levels(DF$Species)
species
```
Looks good. Though the white birch are likely gray birch.

## Step 1. Species Richness
Ok. Now let's get overall species richness by site.
```{r}
SRbySite<-DF %>% group_by(Forest) %>% summarise(
   SR = n_distinct(Species)
)
```
Now by site and by plot
```{r}
SRbyPlot<-DF %>% group_by(Forest, PlotID) %>% summarise(
SR = n_distinct(Species)
)
```
Now mean species richness by forest
```{r}
MeanSR<-SRbyPlot %>% group_by(Forest) %>% summarize(
  MeanSR = mean(SR, na.rm = T),
  SDSR =sd(SR, na.rm = T),
  semSR = (sd(SR)/length(SR)),
  N = length(SR)
)
```
## Step 2. Diversity
We'll use the vegan package to calculate diversity
```{r}
library(vegan)
```

To run diversity indices, we need to know, per site, the number of trees of each species.  Then we transpose into a vegan type matrix.

### Create data set
```{r}
Diversity<-DF %>% group_by(Forest, PlotID, Species) %>% summarize(
count = length(Tree)
)
```
Now we need to spread the data across species, so that each cell is the total number of that species for that plot in that forest

```{r}
Div1<-spread(Diversity, Species, count)
```
This looks good, but we need to get rid of the forest name and plot id to run in vegan. 

```{r}
Div2<-Div1[,-(1:2)]
```
Now replace all the NA's with 0
```{r}
Div2[is.na(Div2)]<-0
```
Now we can use vegan
Shannon index:

```{r}
Shannon<-diversity(Div2, index = "shannon")
Simpson<-diversity(Div2, index = "simpson")
InvSimp<-diversity(Div2, index = "invsimpson")
```
Now append the Diversity data set
```{r}
SRbyPlot$Shannon<-Shannon
SRbyPlot$Simpson<-Simpson
SRbyPlot$Inv_Simpson<-InvSimp
```
And now get all the means
```{r}
MeanDiv<-SRbyPlot %>% group_by(Forest) %>% summarize(
  N = length(SR),
  MeanSR = mean(SR, na.rm = T),
  SDSR =sd(SR, na.rm = T),
  semSR = (sd(SR)/length(SR)),
  MeanShan=mean(Shannon, na.rm = T),
  SD_Shan=sd(Shannon, na.rm = T),
  semShan = (sd(Shannon)/length(SR)),
  MeanSimp = mean(Simpson, na.rm = T),
  SD_Simp=sd(Simpson, na.rm = T),
  semSimp = (sd(Simpson)/length(SR)),
  MeanInSimp = mean(Inv_Simpson, na.rm = T),
  SD_InSimp=sd(Inv_Simpson, na.rm = T),
  semInSimp = (sd(Inv_Simpson)/length(SR)),
  
)
```







