---
title: "Importance Analysis"
output: html_notebook
---

Calculating Leslie's importance values  to see if I get the same results she does when I compare with her work.

Because we measured trees per plot but seedlings and saplings per circle, we need to calculate importance of trees vs. seedlings/saplings and then add them together to get a species importance value. 

Trees are per plot: One plot = 400m2, which is the same as 1/25 of a hectare; thus multiplying by a factor of 25 converts to trees per hectare.

Seedlings/saplings are per circle; one circle is 10m2 which is 1/1000 of a hectare; multiplying by 1000 gives per hectare. However, we have 3 circles per plot; thus we have 30m2 total and that means multiply by 333.3 to get per hectare.

I will get these densities per hectare for calculating importance value.

Prep the workspace
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
```

Start with one file. It has Leslie's data plus the data from the relevant FE forests.

```{r}
DF<-read.csv("Data/forests.csv")
```

Plot some data and look for issues
```{r}
ggplot(DF, aes(DBH))+
   geom_histogram()
```
Looks like there might be a really big tree or two.  Check them out. 

```{r}
big<-filter(DF, DBH >= 60)
big
```
Two big sugar maples at donnerville.  Reasonable.

Now look at species to see if they are all named correctly. Start by making some factor variables
```{r}
species<-levels(DF$Species)
species
```
Looks good. Though the white birch are likely gray birch.

Get a tree and seed/sap dataframe
```{r}
trees<-filter(DF, Type == "tree")
seedsap<-filter(DF, Type != "tree")
```
We will now step through calculation of IV, first for trees and then repeat the same steps for seedlings/saplings.  Then we will do a "combined" for trees+seedlings/saplings.
##Trees
###Step 1. Density and relative density
Establish area var for trees

```{r}
APP_t<-400 #AreaPerPlot = 400m2 per forest plot
```
Determine number of trees per plot
```{r}
PerP<-trees %>% group_by(Forest, PlotID) %>% summarise(
  treesPerP = length(Species) 
)
```
Get total density of trees per plot = #trees/area sampled
```{r}
PerP<-PerP %>% mutate(
  TreesPerM2 = treesPerP/APP_t #units is trees per m2
)
```
Get total density of trees per hectare per forest plot by multiplying DensPerPlot times 10000 (the number of meters squared in a hectare)
```{r}
PerP<-PerP %>% mutate(
  TreesPerHect=TreesPerM2*10000
)
```
Calculate density for each species where density = #trees per area

Start with number of trees by species for each plot
```{r}
ImportancePerPlot<-trees %>% group_by(Forest, PlotID, Species) %>% summarise(
  Number = length(Species),
  NumPerM2 = Number/APP_t,
  NumPerHect = NumPerM2*10000
)
```

Now add the total density per plot to the ImportancePerPlot df so that we can do math with it. 

Use a join to join by forest and plot id and then join PerP to ImportancePerPlot.

```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, PerP, by = c("Forest", "PlotID"))
```
Now I can calculate relative density for each species 

Relative density = (density for a species/total density of trees in plot)*100

```{r}
ImportancePerPlot <- ImportancePerPlot %>% mutate(
  RelDen = (NumPerHect/TreesPerHect)*100 #units are trees of that species per hectare
)
```
Output the overall average tree density for Kate
```{r}
ForestTraits<- ImportancePerPlot %>% group_by(Forest) %>% summarize(
  meanTDensity = mean(TreesPerHect, na.rm = T),
  sdTDensity = sd(TreesPerHect, na.rm = T)
)
```

### Step 2. Dominance and Relative Dominance
Begin by calculating dominance for each species

Dominance = (total of basal area/area sampled)

Calculate basal area for each tree in the trees df
To do so, calculate radius as DBH/2 and then basal area as pi*radius squared

```{r}
trees<-trees %>% mutate(
  BA = (pi*((DBH/2)*(DBH/2)))
)

```
Now get total basal area and dominance (BA/area sampled) per plot

```{r}
Dominance<-trees %>% group_by(Forest, PlotID) %>% summarize(
  TotBA = sum(BA), #units is basal area of all species per plot
  Dom = sum(BA)/APP_t, #units is area of tree per m2 of plot
  DomPerHect = Dom*10000 #units is basal area per hectare of forest
)
```
Now add to ForestTraits for Kate
```{r}
DomSummary<-Dominance %>% group_by(Forest) %>% summarize(
  meanDom = mean(DomPerHect, na.rm = T),
  sdDom = sd(DomPerHect, na.rm = T)
)

ForestTraits$meanTDom<-DomSummary$meanDom
ForestTraits$sdTDom<-DomSummary$sdDom
```

Now add to PerPlot to keep things straightforward
```{r}
PerP$TotBA = Dominance$TotBA
PerP$TotDom=Dominance$Dom
```
Now get BA and dominance per species rather than per plot
```{r}
DomPerSp<-trees %>% group_by(Forest, PlotID, Species) %>% summarize(
  BA = sum(BA),
  Dom = sum(BA)/APP_t
)
```
Now join those to ImportancePerPlot
```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, DomPerSp)
```
Now join the total dominance to ImportancePerP
```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, PerP, by = c("Forest", "PlotID"))
```

Now I can calculate relative dominance

RelDom = (dom for a species/total dom)*100
```{r}
ImportancePerPlot<-ImportancePerPlot %>% mutate(
  RelDom = 100*(Dom/TotDom)
)
```
### Step 3. Frequency and Relative Frequency
First step is to determine the total number of plots. 
```{r}
treePlots<-nrow(PerP)
```
Now determine, for each species, the number of plots in which the species occurs

```{r}
SpeciesByPlots<-trees %>% group_by(Forest, Species, PlotID) %>% summarise(
  NumInPlot = length(PlotID))
#num in plot corresponds to Number in ImportancePerPlot
ImportancePerPlot<-ImportancePerPlot %>% mutate(
  plot_present = 1
)
#assign each species in each plot a value of 1 to indicate present rather than the actual count of trees.

SpByP<-ImportancePerPlot %>% group_by(Species) %>% summarise(
  NumPlots = sum(plot_present)
)

SpByP<-SpByP %>% mutate(
  TotPlots = nrow(PerP),
  RelFreq = (NumPlots/TotPlots)*100
)
  
```
Now join back to importance values
```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, SpByP, by = "Species")
```
### Step 4. Calculate Importance value
Importance value is relative density + relative dominance + relative frequency
```{r}
ImportancePerPlot <-  ImportancePerPlot %>% mutate(
  IV = RelDen+RelDom+RelFreq
)
```
Write these data to a file.

First select just the columns I want

```{r}
TreeIV<-select(ImportancePerPlot, c(Forest, PlotID, Species,RelDen, RelDom,RelFreq, IV))

write.csv(TreeIV, "Data/TreeImportancValues.csv")
```
Now that I have importance value per plot, I want to get average importance value of each species per forest.

```{r}
AvgIVtrees<-ImportancePerPlot %>% group_by(Forest, Species) %>% summarize(
  IV_mean=mean(IV,na.rm=T),
  IV_sd =sd(IV, na.rm = T),
  IV_sem = sd(IV, na.rm = T)/sqrt(sum(!is.na(IV)))
)
```
Now make some plots:

Set up a dodge so that error bars and points won't overlap
```{r}
dodge<-position_dodge(0.35)
```

```{r}
ggplot(AvgIVtrees, aes(Species, IV_mean, group = Forest, color = Forest))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()
```
Now plot by forest rather than species:
```{r}
ggplot(AvgIVtrees, aes(Forest, IV_mean, group = Species, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem), position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()
```
And now plot by species, facet by forest, rank IV first

```{r}
ggplot(AvgIVtrees, aes(Species, IV_mean, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()+
  facet_wrap(~Forest)
```
And finally, facet by species (just to look for patterns)

```{r}
ggplot(AvgIVtrees, aes(Forest, IV_mean, color = Forest))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value - trees")+
  coord_flip()+
  facet_wrap(~Species)+
  theme(legend.position="none")
```
##Seedlings/Saplings
Establish area var for seedlings/saplings
```{r}
APP_ss<-30 #30m2 for the 3 seedling/sapling plots
```

###Step 1. Density and relative density
Determine number of seedlings/saplings per plot
```{r}
sPerP<-seedsap %>% group_by(Forest, PlotID) %>% summarise(
  ssapPerP = length(Species)
)
```
Get density of seedlings and sapling per plot = #seedlings&saplings/area sampled
```{r}
sPerP<-sPerP %>% mutate(
  total_density = ssapPerP/APP_ss,#units is ssap per m2
  ssDensPerHectare = total_density*10000
)
```
Output to forest traits for Kate
```{r}
seedsap_means<-sPerP %>% group_by(Forest) %>% summarize(
  mean_ss_dens_per_hect = mean(ssDensPerHectare, na.rm = T),
  sd_ss_dens_per_hect = sd(ssDensPerHectare, na.rm = T)
)

ForestTraits$mean_ss_dens_per_hect = seedsap_means$mean_ss_dens_per_hect
ForestTraits$sdSSDensity = seedsap_means$sd_ss_dens_per_hect
```


Calculate density for each species where density = #seedsor saps per area

Start with number of seedlings&saplings by species for each plot
```{r}
sImportancePerPlot<-seedsap %>% group_by(Forest, PlotID, Species) %>% summarise(
  Number = length(Species)
)
```
And now get density by species
```{r}
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  dens_in_plot = Number/APP_ss, #units is seedlings/saplings per m2
  dens_per_hect = dens_in_plot*10000 #units is seedlings/saplings per hectare
)
```

Now add the total density per plot to the ImportancePerPlot df so that we can do math with it. 

Use a join to join by forest and plot id and then join PerP to ImportancePerPlot.

```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sPerP, by = c("Forest", "PlotID"))
```

Now I can calculate relative density for each species 

Relative density = (density for a species/total density of trees in plot)*100
```{r}
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  RelDen = (dens_in_plot/total_density)*100
)
```
### Step 2. Dominance and Relative Dominance
Begin by calculating dominance for all seedlings/saplings

Dominance = (total of basal area/area sampled)

Calculate basal area for each tree in the trees df
To do so, calculate radius as DBH/2 and then basal area as pi*radius squared

```{r}
seedsap<-seedsap %>% mutate(
  BA = (pi*((DBH/2)*(DBH/2)))
)

```
Now get total basal area and dominance (BA/area sampled) per plot

```{r}
sDominance<-seedsap %>% group_by(Forest, PlotID) %>% summarize(
  TotBA = sum(BA),
  Dom = sum(BA)/APP_ss, #units is area of seedling/sapling per m2
  DomPerHect = Dom*10000 #units is area of seedlings/saplings per hectare
)
```
Now get averages per forest for Kate's data
```{r}
seedsap_means1<-sDominance %>% group_by(Forest) %>% summarize(
  meanSSDomPerHect = mean(DomPerHect, na.rm = T),
  sdSSDomPerHect = sd(DomPerHect, na.rm = T)
)

ForestTraits$meanSSDom<-seedsap_means1$meanSSDomPerHect
ForestTraits$sdSSDom<-seedsap_means1$sdSSDomPerHect
```
Now write file for Kate's data
```{r}
write.csv(ForestTraits, "Data/ForestTraits.csv")
```

Now add to sPerPlot to keep things straightforward
```{r}
sPerP$TotBA = sDominance$TotBA
sPerP$TotDom=sDominance$Dom
```
Now get BA and dominance per species rather than per plot
```{r}
sDomPerSp<-seedsap %>% group_by(Forest, PlotID, Species) %>% summarize(
  BA = sum(BA),
  Dom = sum(BA)/APP_ss
)
```
Now join those to ImportancePerPlot
```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sDomPerSp)
```
Now join the total dominance to ImportancePerP
```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sPerP, by = c("Forest", "PlotID"))
```

Now I can calculate relative dominance

RelDom = (dom for a species/total dom)*100
```{r}
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  RelDom = 100*(Dom/TotDom)
)
```
### Step 3. Frequency and Relative Frequency
First step is to determine the total number of plots. 
```{r}
seedsapPlots<-nrow(sPerP)
```
Now determine, for each species, the number of plots in which the species occurs

```{r}
sSpeciesByPlots<-seedsap %>% group_by(Forest, Species, PlotID) %>% summarise(
  NumInPlot = length(PlotID))
#num in plot corresponds to Number in ImportancePerPlot
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  plot_present = 1
)
#assign each species in each plot a value of 1 to indicate present rather than the actual count of trees.

sSpByP<-sImportancePerPlot %>% group_by(Species) %>% summarise(
  NumPlots = sum(plot_present)
)

sSpByP<-sSpByP %>% mutate(
  TotPlots = nrow(PerP),
  RelFreq = (NumPlots/TotPlots)*100
)
  
```
Now join back to importance values
```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sSpByP, by = "Species")
```
### Step 4. Calculate Importance value
Importance value is relative density + relative dominance + relative frequency
```{r}
sImportancePerPlot <-  sImportancePerPlot %>% mutate(
  IV = RelDen+RelDom+RelFreq
)
```
Now write to a file. First get just the columns I want.

```{r}
SSAP_IV<-select(sImportancePerPlot, c(Forest, PlotID, Species, RelDen, RelDom, RelFreq, IV))

write.csv(SSAP_IV, "Data/SeedSapImportanceValues.csv")
```


Now that I have importance value per plot, I want to get average importance value of each species per forest.

```{r}
AvgIVseedsap<-sImportancePerPlot %>% group_by(Forest, Species) %>% summarize(
  IV_mean=mean(IV,na.rm=T),
  IV_sd =sd(IV, na.rm = T),
  IV_sem = sd(IV, na.rm = T)/sqrt(sum(!is.na(IV)))
)
```
Now make some plots:
```{r}
ggplot(AvgIVseedsap, aes(Species, IV_mean, group = Forest, color = Forest))+
  geom_point(position = "dodge")+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem), position = "dodge")+
  labs(y = "Average Importance Value - Seedlings/Saplings")+
  coord_flip()
```
Now plot by forest rather than species:
```{r}
ggplot(AvgIVseedsap, aes(Forest, IV_mean, group = Species, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Seedlings/Saplings")+
  coord_flip()
```
And now plot by species, facet by forest, rank IV first

```{r}
ggplot(AvgIVseedsap, aes(Species, IV_mean, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Seedlings/Saplings")+
  coord_flip()+
  facet_wrap(~Forest)
```
And finally, facet by species (just to look for patterns)

```{r}
ggplot(AvgIVseedsap, aes(Forest, IV_mean, color = Forest))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value - seedlings & saplings")+
  coord_flip()+
  facet_wrap(~Species)+
  theme(legend.position="none")
```

## Combine trees with seedlings and saplings
To get overall importance of each species, need to now add together the importance values for trees and seedlings/saplings for each species to give total importance.  Look to the Rmd file `ImportanceAll.Rmd` for those calculations.

