---
title: "Importance Analysis"
output: html_notebook
---

Calculating Leslie's importance values  to see if I get the same results she does when I compare with her work.

This is calcualating importance values per forest, without regard to trees vs. seedlings/saplings (that is, joining seedlings/saplings with trees before calculating).

Trees are per plot: One plot = 400m2, which is the same as 1/25 of a hectare; thus multiplying by a factor of 25 converts to trees per hectare.

Seedlings/saplings are per circle; one circle is 10m2 which is 1/1000 of a hectare; multiplying by 1000 gives per hectare. However, we have 3 circles per plot; thus we have 30m2 total and that means multiply by 333.3 to get per hectare.

I will get these densities per hectare for calculating importance value.

Prep the workspace
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
```

Start with one file. It has Leslie's data plus the data from the relevant FE forests.

```{r}
DF<-read.csv("Data/forests.csv")
```

We will now step through calculation of IV across trees and seedlings/saplings

###Step 1. Density and relative density
Establish area var for trees and for seedlings/saplings

```{r}
APP_t<-400 #AreaPerPlot = 400m2 per forest plot
APP_ss<-30 #Area per plot = 30m2 10m2 circles = 30 m2
circ_to_plot<-13.333 #num sets of 3 seed/sap circles in a plot, i.e. 400/30
```
Determine number of trees per plot

First add a new coding variable of tree vs. seed_sap
```{r}
DF$Code<-NA
trees<-which(DF$Type == "tree")
DF$Code[trees]<-"tree"
seedsap<-which(DF$Type == "seedling" | DF$Type == "sapling")
DF$Code[seedsap]<- "seed_sap"
```
Since seedlings and saplings are per 30m2, but trees are per 400 m2, we need to account for that difference when calculating density.  Get both densities per m2 and then scale to per hectare.  

Now get trees, seedling/saplings by plot
```{r}
PerP<-DF %>% group_by(Forest, Code, PlotID) %>% summarise(
  NumPerPlot = length(Species)
)

trees<-which(PerP$Code == "tree")
seedsap<-which(PerP$Code == "seed_sap")

PerP$NumPerm2<-NA
PerP$NumPerm2[trees]<-PerP$NumPerPlot[trees]/APP_t #units are per m2
PerP$NumPerm2[seedsap]<- PerP$NumPerPlot[seedsap]/APP_ss #units are per m2


PerP$DensPerHect<-NA
PerP$DensPerHect[trees]<-PerP$NumPerm2[trees]*10000 #units are per hectare
PerP$DensPerHect[seedsap]<-PerP$NumPerm2[seedsap]*10000 #units are per hectare
```
Now get total density per hectare of trees and seedlings/saplings by adding together.

```{r}
Density<-PerP %>% group_by(Forest, PlotID) %>% summarize(
  sumDensity = sum(DensPerHect)
)
```
Now join to PerP
```{r}
PerP<-left_join(PerP, Density, by = c("Forest", "PlotID"))
```

Open ForestTraits to add data to spreadsheet
```{r}
ForestTraits<-read.csv("Data/ForestTraits.csv")
```

Now get mean density of trees+seedlings saplings for reporting out to Kate.

```{r}
meanDens<-PerP %>% group_by(Forest) %>% summarize(
  meanDensPerHectAll = mean(sumDensity, na.rm = T),
  sdDensPerHectAll = sd(sumDensity, na.rm = T)
)
```
And add to Forest Traits
```{r}
ForestTraits$MeanAllDens<-meanDens$meanDensPerHectAll
ForestTraits$sdAllDens<-meanDens$sdDensPerHectAll
```
Calculate density for each species where density = #trees or seedlings/saplings per area

Start with number of trees by species for each plot
```{r}
ImportancePerPlot<-DF %>% group_by(Forest, Code, PlotID, Species) %>% summarise(
  Number = length(Species)
)
```
And now get density by species
```{r}
trees<-which(ImportancePerPlot$Code == "tree")
seedsap<-which(ImportancePerPlot$Code == "seed_sap")

ImportancePerPlot$NumPerM2<-NA
ImportancePerPlot$NumPerM2[trees]<-ImportancePerPlot$Number[trees]/APP_t
ImportancePerPlot$NumPerM2[seedsap]<-ImportancePerPlot$Number[seedsap]/APP_ss

ImportancePerPlot$NumPerHect<-NA
ImportancePerPlot$NumPerHect[trees]<-ImportancePerPlot$NumPerM2[trees]*10000
ImportancePerPlot$NumPerHect[seedsap]<-ImportancePerPlot$NumPerM2[seedsap]*10000

```

Now add the total density per hectare to the ImportancePerPlot df so that we can do math with it. 

Use a join to join by forest and plot id and then join PerP to ImportancePerPlot.

```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, Density, by = c("Forest", "PlotID"))
```
Now I can calculate relative density for each species 

Relative density = (density for a species/total density of trees in plot)*100. Need to join sum density from Density to have proper denominator.
```{r}
IPP<-ImportancePerPlot %>% group_by(Forest, PlotID, Species) %>% summarize(
  RelDenAll = sum(RelDenAll, na.rm = T)
)
```
Looks good.

### Step 2. Dominance and Relative Dominance
Begin by calculating dominance for each species

Dominance = (total of basal area/area sampled)

Calculate basal area for each tree and seedling/sapling in the DF
To do so, calculate radius as DBH/2 and then basal area as pi*radius squared

```{r}
DF<-DF %>% mutate(
  BA = (pi*((DBH/2)*(DBH/2))) #units is cm2
)

```
Now get total basal area and dominance (BA/area sampled) per plot

```{r}

DF$Area_sampled<-NA
trees<-which(DF$Code == "tree")
seedsap<-which(DF$Code == "seed_sap")
DF$Area_sampled[trees]<-APP_t
DF$Area_sampled[seedsap]<-APP_ss

DF$BAperM2<- DF$BA/DF$Area_sampled #units are cm2 tree/seedling/sapling per m2 forest
```
Now extract dominance for the forest

```{r}
Dominance<-DF %>% group_by(Forest, PlotID) %>% summarize(
  TotBA = sum(BA, na.rm = T), #units is basal area of all species per plot
  Dom = sum(BAperM2, na.rm = T), #units is area of tree per m2 of plot
  DomPerHect = Dom*10000 #units is basal area per hectare of forest
)
```
Check by calculating separately for trees, seedsap then combine
Trees
```{r}
Trees<-filter(DF, Code == "tree")
```

```{r}
TDominance<-Trees %>% group_by(Forest, PlotID) %>% summarize(
  TotBA = sum(BA), #units is basal area of all species per plot
  Dom = sum(BA)/APP_t, #units is area of tree per m2 of plot
  DomPerHect = Dom*10000 #units is basal area per hectare of forest
)
```
Seedsap
```{r}
SS<-filter(DF, Code == "seed_sap")
```
```{r}
SDominance<-SS %>% group_by(Forest, PlotID) %>% summarize(
  SSTotBA = sum(BA), #units is basal area of all species per plot
  SSDom = sum(BA)/APP_ss, #units is area of tree per m2 of plot
  SSDomPerHect = SSDom*10000 #units is basal area per hectare of forest
)
```
Then rejoin
```{r}
totDom<-TDominance$DomPerHect + SDominance$SSDomPerHect
```
Numbers match what we got using just DF.  Good math check.

Now get averages for Kate
```{r}
meanDom<-Dominance %>% group_by(Forest) %>% summarize(
  meanDominanceAll = mean(Dom, na.rm = T),
  sdDominanceAll = sd(Dom,na.rm = T)
)
```
And now add to ForestTraits
```{r}
ForestTraits$meanAllDom<-meanDom$meanDominanceAll
ForestTraits$sdAllDom<-meanDom$sdDominanceAll
```
And write for Kate
```{r}
write.csv(ForestTraits, "Data/completeForestTraits.csv")
```

Now add to PerPlot to keep things straightforward
```{r}
PerP<-left_join(PerP, Dominance, by = c("Forest", "PlotID"))
```
Now get BA and dominance per species rather than per plot
```{r}

DomPerSp<-DF %>% group_by(Forest, PlotID, Species) %>% summarize(
  TotBA = sum(BA, na.rm = T),
  Dom = sum(BAperM2, na.rm = T),
  DomPerHect = Dom*10000
  )

```
Now bring over total dominance
```{r}
DomPerSp<-left_join(DomPerSp, Dominance, by = c("Forest", "PlotID"))
```
Now I can calculate relative dominance

RelDom = (dom for a species/total dom)*100
```{r}
DomPerSp<-DomPerSp %>% mutate(
  RelDom = 100*(DomPerHect.x/DomPerHect.y)
)
```

Good!

Now join to IPP for calculating importance value

```{r}
IPP$RelDomAll<-DomPerSp$RelDom
```

### Step 3. Frequency and Relative Frequency
First step is to determine the total number of plots. 
```{r}
AllPlots<-length(unique(PerP$PlotID))
```
Now determine, for each species, the number of plots in which the species occurs

```{r}

#num in plot corresponds to Number in ImportancePerPlot
IPP<-IPP %>% mutate(
  plot_present = 1
)
#assign each species in each plot a value of 1 to indicate present rather than the actual count of trees.

SpByP<-IPP %>% group_by(Species) %>% summarise(
  NumPlots = sum(plot_present)
)

SpByP<-SpByP %>% mutate(
  RelFreq = (NumPlots/AllPlots)*100
)
  
```
Now join back to importance values
```{r}
IPP<-left_join(IPP, SpByP, by = "Species")
```

Write output

```{r}
write.csv(IPP, "Data/ForestImportanceValues.csv")
```

### Step 4. Calculate Importance value
Importance value is relative density + relative dominance + relative frequency
```{r}
IPP <-  IPP %>% mutate(
  IV = RelDenAll+RelDomAll+RelFreq
)
```
Now that I have importance value per plot, I want to get average importance value of each species per forest.

```{r}
AvgIV<-IPP %>% group_by(Forest, Species) %>% summarize(
  IV_mean=mean(IV,na.rm=T),
  IV_sd =sd(IV, na.rm = T),
  IV_sem = sd(IV, na.rm = T)/sqrt(sum(!is.na(IV)))
)
```
Now make some plots:

Set up a dodge so that error bars and points won't overlap
```{r}
dodge<-position_dodge(0.35)
```

```{r}
ggplot(AvgIV, aes(Species, IV_mean, group = Forest, color = Forest))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()
```
Now plot by forest rather than species:
```{r}
ggplot(AvgIV, aes(Forest, IV_mean, group = Species, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem), position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()
```
And now plot by species, facet by forest, rank IV first

```{r}
ggplot(AvgIV, aes(Species, IV_mean, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()+
  facet_wrap(~Forest)
```
And finally, facet by species (just to look for patterns)

```{r}
ggplot(AvgIV, aes(Forest, IV_mean, color = Forest))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value - trees")+
  coord_flip()+
  facet_wrap(~Species)+
  theme(legend.position="none")
```
