---
title: "Importance Analysis"
output: html_notebook
---

Calculating Leslie's importance values  to see if I get the same results she does when I compare with her work.

Because we measured trees per plot but seedlings and saplings per circle, we need to calculate importance of trees vs. seedlings/saplings and then add them together to get a species importance value. 

Trees are per plot: One plot = 400m2, which is the same as 1/25 of a hectare; thus multiplying by a factor of 25 converts to trees per hectare.

Seedlings/saplings are per circle; one circle is 10m2 which is 1/1000 of a hectare; multiplying by 1000 gives per hectare. However, we have 3 circles per plot; thus we have 30m2 total and that means multiply by 333.3 to get per hectare.

Prep the workspace
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
```

Start with one file. It has Leslie's data plus the data from the relevant FE forests.

```{r}
DF<-read.csv("Data/forests.csv")
```

Plot some data and look for issues
```{r}
ggplot(DF, aes(DBH))+
   geom_histogram()
```
Looks like there might be a really big tree or two.  Check them out. 

```{r}
big<-filter(DF, DBH >= 60)
big
```
Two big sugar maples at donnerville.  Reasonable.

Now look at species to see if they are all named correctly. Start by making some factor variables
```{r}
species<-levels(DF$Species)
species
```
Looks good. Though the white birch are likely gray birch.

Get a tree and seed/sap dataframe
```{r}
trees<-filter(DF, Type == "tree")
seedsap<-filter(DF, Type != "tree")
```
We will now step through calculation of IV, first for trees and then repeat the same steps for seedlings/saplings.
##Trees
###Step 1. Density and relative density
Establish area var for trees

```{r}
APP_t<-400 #AreaPerPlot = 400m2 per forest plot
```
Determine number of trees per plot
```{r}
PerP<-trees %>% group_by(Forest, PlotID) %>% summarise(
  treesPerP = length(Species)
)
```
Get total density of trees per plot = #trees/area sampled
```{r}
PerP<-PerP %>% mutate(
  DensPerPlot = treesPerP/APP_t
)
```
Get total density of trees per hectare per forest plot by multiplying DensPerPlot times 25 (the scaling factor for a forest plot)
```{r}
PerP<-PerP %>% mutate(
  DensPerHect=DensPerPlot*25
)
```
Calculate density for each species where density = #trees per area

Start with number of trees by species for each plot
```{r}
ImportancePerPlot<-trees %>% group_by(Forest, PlotID, Species) %>% summarise(
  Number = length(Species)
)
```
And now get density by species
```{r}
ImportancePerPlot<-ImportancePerPlot %>% mutate(
  PlotDensity = Number/APP_t
)
```
Now add the total density per plot to the ImportancePerPlot df so that we can do math with it. 

Use a join to join by forest and plot id and then join PerP to ImportancePerPlot.

```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, PerP, by = c("Forest", "PlotID"))
```
It worked! Now drop a couple of columns and rename a couple to keep things making sense:
```{r}
#drop density per hectare
ImportancePerPlot<-select(ImportancePerPlot, -c(treesPerP,DensPerHect))
#now rename
ImportancePerPlot<-rename(ImportancePerPlot, dens_in_plot = PlotDensity, total_density = DensPerPlot )
```
Now I can calculate relative density for each species 

Relative density = (density for a species/total density of trees in plot)*100
```{r}
ImportancePerPlot<-ImportancePerPlot %>% mutate(
  RelDen = (dens_in_plot/total_density)*100
)
```
### Step 2. Dominance and Relative Dominance
Begin by calculating dominance for each species

Dominance = (total of basal area/area sampled)

Calculate basal area for each tree in the trees df
To do so, calculate radius as DBH/2 and then basal area as pi*radius squared

```{r}
trees<-trees %>% mutate(
  BA = (pi*((DBH/2)*(DBH/2)))
)

```
Now get total basal area and dominance (BA/area sampled) per plot

```{r}
Dominance<-trees %>% group_by(Forest, PlotID) %>% summarize(
  TotBA = sum(BA),
  Dom = sum(BA)/APP_t
)
```
Now add to PerPlot to keep things straightforward
```{r}
PerP$TotBA = Dominance$TotBA
PerP$TotDom=Dominance$Dom
```
Now get BA and dominance per species rather than per plot
```{r}
DomPerSp<-trees %>% group_by(Forest, PlotID, Species) %>% summarize(
  BA = sum(BA),
  Dom = sum(BA)/APP_t
)
```
Now join those to ImportancePerPlot
```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, DomPerSp)
```
Now join the total dominance to ImportancePerP
```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, PerP, by = c("Forest", "PlotID"))
```
Now drop a couple of columns and rename a couple to keep things making sense:
```{r}
#drop unneeded variables
ImportancePerPlot<-select(ImportancePerPlot, -c(treesPerP,DensPerPlot, DensPerHect, TotBA))

```
Now I can calculate relative dominance

RelDom = (dom for a species/total dom)*100
```{r}
ImportancePerPlot<-ImportancePerPlot %>% mutate(
  RelDom = 100*(Dom/TotDom)
)
```
### Step 3. Frequency and Relative Frequency
First step is to determine the total number of plots. 
```{r}
treePlots<-nrow(PerP)
```
Now determine, for each species, the number of plots in which the species occurs

```{r}
SpeciesByPlots<-trees %>% group_by(Forest, Species, PlotID) %>% summarise(
  NumInPlot = length(PlotID))
#num in plot corresponds to Number in ImportancePerPlot
ImportancePerPlot<-ImportancePerPlot %>% mutate(
  plot_present = 1
)
#assign each species in each plot a value of 1 to indicate present rather than the actual count of trees.

SpByP<-ImportancePerPlot %>% group_by(Species) %>% summarise(
  NumPlots = sum(plot_present)
)

SpByP<-SpByP %>% mutate(
  TotPlots = nrow(PerP),
  RelFreq = (NumPlots/TotPlots)*100
)
  
```
Now join back to importance values
```{r}
ImportancePerPlot<-left_join(ImportancePerPlot, SpByP, by = "Species")
```
### Step 4. Calculate Importance value
Importance value is relative density + relative dominance + relative frequency
```{r}
ImportancePerPlot <-  ImportancePerPlot %>% mutate(
  IV = RelDen+RelDom+RelFreq
)
```
Now that I have importance value per plot, I want to get average importance value of each species per forest.

```{r}
AvgIVtrees<-ImportancePerPlot %>% group_by(Forest, Species) %>% summarize(
  IV_mean=mean(IV,na.rm=T),
  IV_sd =sd(IV, na.rm = T),
  IV_sem = sd(IV, na.rm = T)/sqrt(sum(!is.na(IV)))
)
```
Now make some plots:

Set up a dodge so that error bars and points won't overlap
```{r}
dodge<-position_dodge(0.35)
```

```{r}
ggplot(AvgIVtrees, aes(Species, IV_mean, group = Forest, color = Forest))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()
```
Now plot by forest rather than species:
```{r}
ggplot(AvgIVtrees, aes(Forest, IV_mean, group = Species, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem), position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()
```
And now plot by species, facet by forest, rank IV first

```{r}
ggplot(AvgIVtrees, aes(Species, IV_mean, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Trees")+
  coord_flip()+
  facet_wrap(~Forest)
```
And finally, facet by species (just to look for patterns)

```{r}
ggplot(AvgIVtrees, aes(Forest, IV_mean, color = Forest))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value - trees")+
  coord_flip()+
  facet_wrap(~Species)+
  theme(legend.position="none")
```
##Seedlings/Saplings
Establish area var for seedlings/saplings
```{r}
APP_ss<-30 #30m2 for the 3 seedling/sapling plots
```

###Step 1. Density and relative density
Determine number of trees per plot
```{r}
sPerP<-seedsap %>% group_by(Forest, PlotID) %>% summarise(
  ssapPerP = length(Species)
)
```
Get total density of seedlings and sapling per plot = #seedlings&saplings/area sampled
```{r}
sPerP<-sPerP %>% mutate(
  total_density = ssapPerP/APP_ss
)
```

Calculate density for each species where density = #seedsor saps per area

Start with number of seedlings&saplings by species for each plot
```{r}
sImportancePerPlot<-seedsap %>% group_by(Forest, PlotID, Species) %>% summarise(
  Number = length(Species)
)
```
And now get density by species
```{r}
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  dens_in_plot = Number/APP_ss
)
```
Now add the total density per plot to the ImportancePerPlot df so that we can do math with it. 

Use a join to join by forest and plot id and then join PerP to ImportancePerPlot.

```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sPerP, by = c("Forest", "PlotID"))
```

Now I can calculate relative density for each species 

Relative density = (density for a species/total density of trees in plot)*100
```{r}
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  RelDen = (dens_in_plot/total_density)*100
)
```
### Step 2. Dominance and Relative Dominance
Begin by calculating dominance for each species

Dominance = (total of basal area/area sampled)

Calculate basal area for each tree in the trees df
To do so, calculate radius as DBH/2 and then basal area as pi*radius squared

```{r}
seedsap<-seedsap %>% mutate(
  BA = (pi*((DBH/2)*(DBH/2)))
)

```
Now get total basal area and dominance (BA/area sampled) per plot

```{r}
sDominance<-seedsap %>% group_by(Forest, PlotID) %>% summarize(
  TotBA = sum(BA),
  Dom = sum(BA)/APP_ss
)
```
Now add to sPerPlot to keep things straightforward
```{r}
sPerP$TotBA = sDominance$TotBA
sPerP$TotDom=sDominance$Dom
```
Now get BA and dominance per species rather than per plot
```{r}
sDomPerSp<-seedsap %>% group_by(Forest, PlotID, Species) %>% summarize(
  BA = sum(BA),
  Dom = sum(BA)/APP_ss
)
```
Now join those to ImportancePerPlot
```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sDomPerSp)
```
Now join the total dominance to ImportancePerP
```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sPerP, by = c("Forest", "PlotID"))
```
Now drop a couple of columns and rename a couple to keep things making sense:

Now I can calculate relative dominance

RelDom = (dom for a species/total dom)*100
```{r}
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  RelDom = 100*(Dom/TotDom)
)
```
### Step 3. Frequency and Relative Frequency
First step is to determine the total number of plots. 
```{r}
seedsapPlots<-nrow(sPerP)
```
Now determine, for each species, the number of plots in which the species occurs

```{r}
sSpeciesByPlots<-seedsap %>% group_by(Forest, Species, PlotID) %>% summarise(
  NumInPlot = length(PlotID))
#num in plot corresponds to Number in ImportancePerPlot
sImportancePerPlot<-sImportancePerPlot %>% mutate(
  plot_present = 1
)
#assign each species in each plot a value of 1 to indicate present rather than the actual count of trees.

sSpByP<-sImportancePerPlot %>% group_by(Species) %>% summarise(
  NumPlots = sum(plot_present)
)

sSpByP<-sSpByP %>% mutate(
  TotPlots = nrow(PerP),
  RelFreq = (NumPlots/TotPlots)*100
)
  
```
Now join back to importance values
```{r}
sImportancePerPlot<-left_join(sImportancePerPlot, sSpByP, by = "Species")
```
### Step 4. Calculate Importance value
Importance value is relative density + relative dominance + relative frequency
```{r}
sImportancePerPlot <-  sImportancePerPlot %>% mutate(
  IV = RelDen+RelDom+RelFreq
)
```
Now that I have importance value per plot, I want to get average importance value of each species per forest.

```{r}
AvgIVseedsap<-sImportancePerPlot %>% group_by(Forest, Species) %>% summarize(
  IV_mean=mean(IV,na.rm=T),
  IV_sd =sd(IV, na.rm = T),
  IV_sem = sd(IV, na.rm = T)/sqrt(sum(!is.na(IV)))
)
```
Now make some plots:
```{r}
ggplot(AvgIVseedsap, aes(Species, IV_mean, group = Forest, color = Forest))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Seedlings/Saplings")+
  coord_flip()
```
Now plot by forest rather than species:
```{r}
ggplot(AvgIVseedsap, aes(Forest, IV_mean, group = Species, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Seedlings/Saplings")+
  coord_flip()
```
And now plot by species, facet by forest, rank IV first

```{r}
ggplot(AvgIVseedsap, aes(Species, IV_mean, color = Species))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem),position = dodge)+
  labs(y = "Average Importance Value - Seedlings/Saplings")+
  coord_flip()+
  facet_wrap(~Forest)
```
And finally, facet by species (just to look for patterns)

```{r}
ggplot(AvgIVseedsap, aes(Forest, IV_mean, color = Forest))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value - seedlings & saplings")+
  coord_flip()+
  facet_wrap(~Species)+
  theme(legend.position="none")
```

## Combine trees with seedlings and saplings
To get overall importance of each species, need to now add together the importance values for trees and seedlings/saplings for each species to give total importance.  

I just made this up as an approach.

Do this by first joining the tables for trees and for seedlings/saplings:

```{r}
allImportance<-full_join(ImportancePerPlot, sImportancePerPlot, by = c("Forest", "PlotID", "Species"))

```
Now strip down to just the columns I need:
```{r}
keep<-c("Forest", "PlotID", "Species", "IV.x", "IV.y")
Imp<-allImportance[,keep]
```
Now replace NAs with zeros for IVs for those species that are not present at certain sites.

```{r}
Imp[is.na(Imp)]<-0
```

Now add together, then get average per forest
```{r}
Imp<-Imp %>% mutate(
  sumIV = IV.x+ IV.y
  )#note that this will be problematic if you have NA's
```
Now get averages
```{r}
AvgIV<-Imp %>% group_by(Forest, Species) %>% summarize(
  IV_mean=mean(sumIV,na.rm=T),
  IV_sd =sd(sumIV, na.rm = T),
  IV_sem = sd(sumIV, na.rm = T)/sqrt(sum(!is.na(sumIV)))
)
```
And plot:
Set up a dodge to keep things from plotting on top of one another:
```{r}
dodge<-position_dodge(0.35)
```

```{r}
ggplot(AvgIV, aes(Species, IV_mean, group = Forest, color = Forest))+
  geom_point(position = dodge)+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem), position = dodge)+
  labs(y = "Average Importance Value")+
  coord_flip()
```


Now plot by forest rather than species:
```{r}
ggplot(AvgIV, aes(Forest, IV_mean, group = Species, color = Species))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value")+
  coord_flip()
```
And now plot by species, facet by forest, rank IV first

```{r}
ggplot(AvgIV, aes(Species, IV_mean, color = Species))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value")+
  coord_flip()+
  facet_wrap(~Forest)+
  theme(legend.position="none")
```
And finally, facet by species (just to look for patterns)

```{r}
ggplot(AvgIV, aes(Forest, IV_mean, color = Forest))+
  geom_point()+
  geom_errorbar(aes(ymin = IV_mean - IV_sem, ymax = IV_mean + IV_sem))+
  labs(y = "Average Importance Value")+
  coord_flip()+
  facet_wrap(~Species)+
  theme(legend.position="none")
```
Finally, let's generate a table of imporatance values by forest, plot and species.We'll include tree IV, seed/sap IV, and total IV.

These are the data contained in Imp, but let's rename the columns for clarity before writing it.

```{r}
names(Imp)
```
Now rename
```{r}
Imp <- Imp %>% rename(IV_tree = IV.x, IV_ssap = IV.y, IV_all = sumIV)
```
And write
```{r}
write.csv(Imp,"Data/ImportanceValues.csv")
```




