---
title: "R Notebook"
output: html_notebook
---

Let's do a community ordination on the forests from the Game Camera study and see how the forests compare to each other.  Use code from Forest Ecology as an example.

To do this, we need a data frame that has a column for site, a column for plot and then a column for each species.  The value in each cell for each species should be the value upon with the ordination is being based (e.g. importance value, number of species, etc.)

We'll use output generated from the importance value work we did in Importance.Rmd.

### Set up the workspace
```{r}
rm(list = ls())
library(tidyverse)
library(vegan)
library(ggplot2)
```
Install ggvegan from github
```{r}
devtools::install_github("gavinsimpson/ggvegan")
```
And turn it on
```{r}
library(ggvegan)
```
And load some data

```{r}
IV<-read.csv("Data/ImportanceValues.csv", stringsAsFactors = F)
```
Now let's divide into 3 df's for all, trees, seedsaps

```{r}
IV_all<-select(IV, -c(IV_tree, IV_ssap))
IV_trees<-select(IV, -c(IV_ssap, IV_all))
IV_ssap<-select(IV, -c(IV_tree, IV_all))
```
And now we need to rearrange to fit the standard vegan shape.

```{r}
IVaV<-spread(IV_all, Species, IV_all)
IVtV<-spread(IV_trees, Species, IV_tree)
IVsV<-spread(IV_ssap, Species, IV_ssap)
```
Now need to replace all the NAs with zeroes
```{r}
IVaV[is.na(IVaV)]<-0
IVtV[is.na(IVtV)]<-0
IVsV[is.na(IVsV)]<-0
```

### Run the ordination
#### Start with all
```{r}
IVa.nmds.bray<-metaMDS(IVaV[,3:22], distance = "bray", trace = FALSE, trymax = 100)
```
metaMDS is a "wrapper function" - takes raw species abundance data and standardizes is so that abundance isn't over important.  Then runs
monoMDS a bunch of times until it times 2 similar configurations with minimized stress value and rotates answer.  trymax is number of tries
 
Now look at the results
```{r}
IVa.nmds.bray
```
Plot the result
```{r}
plot(IVa.nmds.bray, type = "t")
```
And make a nicer plot
```{r}
PlotData<-fortify(IVa.nmds.bray)
```
And plot:
```{r}
ggplot(PlotData, aes(NMDS1,NMDS2,group = Score))+
   geom_text(aes(label=Label, color = Score))+
   scale_color_manual(values = c(sites = "brown", species = "green"))+
   geom_point(aes(color=Score, shape =Score, size =.1))+
   scale_color_manual(values = c(sites = "red", species = "black"))
```
Goodness of fit
```{r}
gof<-goodness(IVa.nmds.bray) #goodness of fit
plot(IVa.nmds.bray, type = "t", main = "goodness of fit")
points(IVa.nmds.bray, display = "sites", cex=gof*100)

```
#### Now trees
```{r}
IVt.nmds.bray<-metaMDS(IVtV[,3:22], distance = "bray", trace = FALSE, trymax = 100)
```
metaMDS is a "wrapper function" - takes raw species abundance data and standardizes is so that abundance isn't over important.  Then runs
monoMDS a bunch of times until it times 2 similar configurations with minimized stress value and rotates answer.  trymax is number of tries
 
Now look at the results
```{r}
IVt.nmds.bray
```
Plot the result
```{r}
plot(IVt.nmds.bray, type = "t")
```
And make a nicer plot
```{r}
PlotDataT<-fortify(IVt.nmds.bray)
```
And plot:
```{r}
ggplot(PlotDataT, aes(NMDS1,NMDS2,group = Score))+
   geom_text(aes(label=Label, color = Score))+
   scale_color_manual(values = c(sites = "brown", species = "green"))+
   geom_point(aes(color=Score, shape =Score, size =.1))+
   scale_color_manual(values = c(sites = "red", species = "black"))
```
Goodness of fit
```{r}
gof<-goodness(IVt.nmds.bray) #goodness of fit
plot(IVt.nmds.bray, type = "t", main = "goodness of fit")
points(IVt.nmds.bray, display = "sites", cex=gof*100)

```
#### And seedlings/saplings:
```{r}
IVs.nmds.bray<-metaMDS(IVsV[,3:22], distance = "bray", trace = FALSE, trymax = 100)
```
metaMDS is a "wrapper function" - takes raw species abundance data and standardizes is so that abundance isn't over important.  Then runs
monoMDS a bunch of times until it times 2 similar configurations with minimized stress value and rotates answer.  trymax is number of tries
 
Now look at the results
```{r}
IVs.nmds.bray
```
Plot the result
```{r}
plot(IVs.nmds.bray, type = "t")
```
And make a nicer plot
```{r}
PlotDataS<-fortify(IVs.nmds.bray)
```
And plot:
```{r}
ggplot(PlotDataS, aes(NMDS1,NMDS2,group = Score))+
   geom_text(aes(label=Label, color = Score))+
   scale_color_manual(values = c(sites = "brown", species = "green"))+
   geom_point(aes(color=Score, shape =Score, size =.1))+
   scale_color_manual(values = c(sites = "red", species = "black"))
```
Goodness of fit
```{r}
gof<-goodness(IVs.nmds.bray) #goodness of fit
plot(IVs.nmds.bray, type = "t", main = "goodness of fit")
points(IVs.nmds.bray, display = "sites", cex=gof*100)

```





